---
title: "Getting Started with netOP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with netOP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
library(netOP)
set.seed(2026)
```

## Generate and inspect a network

Generators return adjacency matrices directly and attach only compact truth
metadata. This keeps sparse output useful without attaching a dense probability
matrix.

```{r generate}
A <- generate_sbm(
  n = 30,
  K = 2,
  alpha = 0.5,
  beta = 0.08,
  representation = "dense",
  seed = 1,
  ncores = 1
)
parameters <- get_generator_parameters(A)
table(parameters$g_true)
```

## Embed and cluster

```{r fit}
embedding <- ase(A, d = 2)
dim(embedding$Z_hat)

clustering <- spectral_cluster(
  A,
  K = 2,
  spectral_engine = "base",
  cluster_engine = "kmeans"
)
table(clustering$g_hat)
```

## Select a model

The public model-selection APIs begin with the network and candidate set.
Examples use one worker and small deterministic inputs; production analyses can
increase repetition counts and choose partial eigensolvers.

```{r select}
selection <- netcrop_blockmodel(
  A,
  K_candidates = 1:2,
  num_subnetworks = 2,
  overlap_size = 10,
  nrep = 1,
  losses = "sse",
  ncores = 1,
  seed = 2,
  verbose = FALSE,
  sbm_est_options = list(spectral_cluster = list(spectral_engine = "base")),
  dcbm_est_options = list(spectral_cluster = list(spectral_engine = "base"))
)
selection$best_model_overall
```

NETCROP is also available for RDPG and latent-space dimensions and spectral
regularization. The self-contained ECV and NCV wrappers provide alternative
block-model stability selectors; see `?ecv_stability_blockmodel` and
`?ncv_stability_blockmodel` for disclosures, algorithm restrictions, and
citations.

## Results and plotting

All high-level model-selection results provide `print()` and `summary()`
methods. Plotting is available when `ggplot2` is installed.

```{r inspect}
summary(selection)
if (requireNamespace("ggplot2", quietly = TRUE)) {
  plot(selection)
}
```
